\documentclass{article}
\input{../../.latex/article.tex}
\input{../../.latex/macros.tex}

\newcommand{\assignment}{Section 2 Binary Operations, p25 1,3,5,27,28,36}

\lhead{MAT 311 Abstract Algebra}
\chead{\assignment}
\rhead{Peter Schaefer}

\begin{document}
\section*{\assignment}

Exercises 1 through 4 concern the binary operation $*$ defined on $S = \{a,b,c,d,e\}$ by means of Table 2.26 (not shown).
\begin{enumerate}
    \qitem{1}{Compute $b*d$, $c*c$, and $[(a*c)*e]*a$}{
        Here are the computations:
        \begin{align*}
            b*d & = e \\
            c*c & = b \\
            [(a*c)*e]*a = [c*e]*a = a*a & = a
        \end{align*}
    }
    \qitem{3}{Compute $(b*d)*c \tand b*(d*c)$. Can you say on the basis of these computations whether $*$ is associative?}{
        Examples can only tell us if $*$ is not associative.
        \begin{align*}
            (b*d)*c & = e*c = a \\
            b*(d*c) & = b*b = c
        \end{align*}
        Since $a \neq c$, we know that $*$ is not associative.
    }
    \qitem{5}{Complete Table 2.27 so as to define a commutative binary operation $*$ on $S = \{a,b,c,d\}$.}{
        \bld{2.28 Table}
        \[
            \begin{array}{c|cccc}
                * & a & b & c & d \\
                \hline
                a & a & b & c & \bld{d} \\
                b & b & d & \bld{a} & c \\
                c & c & a & d & b \\
                d & d & \bld{c} & \bld{b} & a
            \end{array}
        \]
    }
\end{enumerate}

In Exercise 27 and 28, either prove the statement or give a counterexample.
\begin{enumerate}
    \qitem{27}{Every binary operation on a set consisting of a single element (is) commutative and associative.}{
        There is only one unique set consisting of a single element.
        \begin{proof}
            Consider $S = \{s\} \twhere s*s = s$.
            \begin{enumerate}
                \item Commutative: $s*s = s = s*s$. Thus $S$ is commutative under $*$.
                \item Associative: $s*(s*s) = s*s = s = s*s = (s*s)*s$. Thus $S$ is associative under $*$.
            \end{enumerate}
            Thus any binary operation on a set consisting of a single element is commutative and associative.
        \end{proof}
    }
    \qitem{28}{Every commutative binary operation on a set having just two elements is associative.}{answer}
\end{enumerate}

\question{36}{Suppose that $*$ is an \itl{associative binary operation} on a set $S$. Let $H = \{a \in S : a * x = x * a ~\text{for all}~ x \in S\}$. Show that $H$ is closed under $*$. (We think of $H$ as consisting of all elements of $S$ that \itl{commute} with every element in $S$.)}

\end{document}