\input{../templates/packages-article.tex}
\input{../templates/macros.tex}

\lhead{Linear Algebra}
\chead{Homework 4}
\rhead{Peter Schaefer}

\begin{document}

\section*{Homework 4}

\subsection*{Problem 9}
Let $V = \bb{R}^3$ be a vector space with standard addition and scalar multiplication. Use Theorem 3 of Lecture Note 8 to determine whether the following sets are subspaces of $V$.
\begin{enumerate}
  \biitem{b}{The set of vectors of the form $(a,1,1)$, $W$}
  \begin{proof}
    Axiom 1: Consider $\vec{a} = (a,1,1) \tand \vec{b} = (b,1,1) \tfor a,b \in \bb{R}$. $\vec{a} \in W \tand \vec{b} \in W$.
    \begin{align*}
      \vec{a} \oplus_W \vec{b} & = (a,1,1) \oplus (b,1,1) = (a+b,1+1,1+1) \\
                               & = (a+b,2,2) \not \in W
    \end{align*}
    Therefore $W$ is \bld{not} closed under addition. \\
    Since Axiom 1 does not hold for $W$, $W$ is not a subspace of $V$.
  \end{proof}
  \biitem{c}{The set of vectors of the form $(a,b,c)$, where $b = a + c$, $W$}
  \begin{proof}
    Axiom 1: Consider $\vec{v} = (a_1,b_1,c_1) \tand \vec{u} = (a_2,b_2,c_2) \tfor a_1,a_2,b_1,b_2,c_1,c_2 \in \bb{R}$. \\
    $\vec{v} \in W \tand \vec{u} \in W$.
    \begin{align*}
      \vec{v} \oplus \vec{u} & = (a_1,b_1,c_1) \oplus (a_2,b_2,c_2) = (a_1+a_2,b_1+b_2,c_1+c_2)   \\
                             & \in W                                                              \\ \\
      b_1 + b_2              & = (a_1 + c_1) + (a_2 + c_2) = (a_1 + a_2) + (c_1 + c_2)~\checkmark
    \end{align*}
    Therefore $W$ is closed under addition. \\
    Axiom 6: Consider $\vec{v} = (a,b,c) \tsuchthat a,b,c \in \bb{R} \tand k \in \bb{R}$. Let $b = a+c$.
    \begin{align*}
      k \odot \vec{v} & = k \odot (a,b,c) = k \odot (a,a+c,c) = (ka, k(a+c), kc) = (ka, ka+kc, kc) \\
                      & \in W                                                                      \\ \\
      kb              & = k(a+c) = ka + kc~\checkmark
    \end{align*}
    Therefore $W$ is closed under scalar multiplication. \\
    Since Axiom 1 and Axiom 6 hold for $W$, $\oplus \tand \odot$ are inherited from $V$, and $W \subseteq V$, through the use of Theorem 3, $W$ is a subspace of $V$.
  \end{proof}
  \biitem{d}{The set of vectors of the form $(a,b,0)$, $W$}
  \begin{proof}
    Axiom 1: Consider $\vec{v} = (a_1,b_1,0) \tand \vec{u} = (a_2,b_2,0) \tfor a_1,a_2,b_1,b_2 \in \bb{R}$.
    \begin{align*}
      \vec{v} \oplus \vec{u} & = (a_1,b_1,0) \oplus (a_2,b_2,0) = (a_1+a_2,b_1+b_2,0+0) = (a_1+a_2,b_1+b_2,0) \\
                             & \in W~\text{since it takes the form}~(a,b,0)
    \end{align*}
    Therefore $W$ is closed under addition. \\
    Axiom 6: Consider $\vec{v} = (a,b,0) \tsuchthat a,b \in \bb{R} \tand k \in \bb{R}$.
    \begin{align*}
      k \odot \vec{v} & = k \odot (a,b,0) = (ka,kb,k0) = (ka,kb,0)   \\
                      & \in W~\text{since it takes the form}~(a,b,0)
    \end{align*}
    Therefore $W$ is closed under scalar multiplication. \\
    Since Axiom 1 and Axiom 6 hold for $W$, $\oplus \tand \odot$ are inherited from $V$, and $W \subseteq V$, through the use of Theorem 3, $W$ is a subspace of $V$.
  \end{proof}
\end{enumerate}

\subsection*{Problem 10}
Let $V = P_3$ be the vector space of all polynomials with degree \itl{up to} 3, with standard addition and scalar multiplication. Use Theorem 3 of Lecture Note 8 to determine whether the following sets are subspaces of $V$.
\begin{enumerate}
  \biitem{b}{The set of polynomials $a_0+a_1x+a_2x^2+a_3x^3$ for which $a_0+a_1+a_2+a_3=0$, $W$.}
  \begin{proof}
    Axiom 1: Consider $\vec{a} = a_0+a_1x+a_2x^2+a_3x^3 \tand \vec{b} = b_0+b_1x+b_2x^2+b_3x^3 \twhere a_{0-3}, b_{0-3} \in \bb{R}$. \\
    Let $a_0+a_1+a_2+a_3=0 \tand b_0+b_1+b_2+b_3=0$.
    \begin{align*}
      \vec{a} \oplus \vec{b}                        & = (a_0+a_1x+a_2x^2+a_3x^3) \oplus (b_0+b_1x+b_2x^2+b_3x^3) \\
                                                    & = a_0+a_1x+a_2x^2+a_3x^3 + b_0+b_1x+b_2x^2+b_3x^3          \\
                                                    & = (a_0+b_0) + (a_1+b_1)x + (a_2+b_2)x^2 + (a_3+b_3)x^3     \\
                                                    & \in W                                                      \\ \\
      (a_0+b_0) + (a_1+b_1) + (a_2+b_2) + (a_3+b_3) & = a_0 + a_1 + a_2 + a_3 + b_0 + b_1 + b_2 + b_3            \\
                                                    & = 0 + 0 = 0~\checkmark
    \end{align*}
    Therefore $W$ is closed under addition. \\
    Axiom 6: Consider $\vec{a} = a_0+a_1x+a_2x^2+a_3x^3 \tsuchthat a_{0-3} \in \bb{R} \tand k \in \bb{R}$. \\
    Let $a_0+a_1+a_2+a_3=0$.
    \begin{align*}
      k \odot \vec{a} & = k \odot (a_0+a_1x+a_2x^2+a_3x^3) = k(a_0+a_1x+a_2x^2+a_3x^3)           \\
                      & = ka_0+ka_1x+ka_2x^2+ka_3x^3                                             \\
                      & \in W                                                                    \\ \\
      ka_0            & +ka_1x+ka_2x^2+ka_3x^3 = k(a_0+a_1x+a_2x^2+a_3x^3) = k(0) = 0~\checkmark
    \end{align*}
    Therefore $W$ is closed under scalar multiplication. \\
    Since Axiom 1 and Axiom 6 hold for $W$, $\oplus \tand \odot$ are inherited from $V$, and $W \subseteq V$, through the use of Theorem 3, $W$ is a subspace of $V$.
  \end{proof}
  \biitem{c}{The set of polynomials $a_0+a_1x+a_2x^2+a_3x^3$ in which $a_0,a_1,a_2, \tand a_3$ are integers.}
  \begin{proof}
    Axiom 6: Consider $\vec{a} = 1+1x+1x^2+1x^3 \tand k = 0.66$.
    \begin{align*}
      k \odot \vec{a} & = 0.66 \odot (1+1x+1x^2+1x^3) = 0.66(1+1x+1x^2+1x^3) \\
                      & = 0.66+0.66x+0.66x^2+0.66x^3                         \\
                      & \not \in W, \tsince 0.66 \not \in \bb{Z}             \\
    \end{align*}
    Therefore $W$ is \bld{not} closed under scalar multiplication. \\
    Since Axiom 6 does not hold for $W$, $W$ is not a subspace of $V$.
  \end{proof}
\end{enumerate}

\subsection*{Problem 11}
Let $V = F(-\infty, \infty)$ be the vector space of all functions from $\bb{R}$ to $\bb{R}$, with standard addition and scalar multiplication. Use Theorem 3 of Lecture Note 8 to determine whether the following sets are subspaces of $V$.
\begin{enumerate}
  \biitem{b}{The set of functions $f$ in $F(-\infty, \infty)$ for which $f(0) = 1$.}
  \begin{proof}
    Axiom 1: Consider $\vec{f}(x) = e^x \tand \vec{g}(x) = e^x$. $\vec{f}, \vec{g} \in W \tsince e^0 = 0$.
    \begin{align*}
      (\vec{f} \oplus \vec{g})(x) & = \vec{f}(x) + \vec{g}(x) = e^x + e^x = 2e^x \not \in W \\
      \twhen x                    & = 0: 2e^0 = 2 \cdot 1 = 2 \neq 1
    \end{align*}
    Therefore $W$ is \bld{not} closed under addition. \\
    Since Axiom 1 does not hold for $W$, $W$ is not a subspace of $V$.
  \end{proof}
  \biitem{c}{The set of functions $f$ in $F(-\infty, \infty)$ for which $f(-x) = x$}
  \begin{proof}

  \end{proof}
\end{enumerate}

\subsection*{Problem 13}
Let $V$ be a vector space. Let $I$ be a nonempty set (often called the "index set"), and let $W_i$ be a subspace of $V$ for all $i \in I$. Prove that $\displaystyle \bigcap_{i \in I} W_i$, is a subspace of V.
\begin{proof}

\end{proof}

\end{document}