\section{Binary Operations}

\subsubsection{Def. Binary Operation}
A \bld{binary operation} $*$ on a set $S$ is a function from $S \times S$ into $S$, $*: S \times S \rightarrow S$. That is, $*$ is a rule which assigns to each ordered pair $(a,b) \in S \times S$ exactly one element $a * b \in S$.

\subsubsection*{Condition 1: Uniquely Defined}
For all $a,b \in S \times S$, $a * b$ must be \bld{uniquely defined}. This means that $*$ cannot be undefined for any $a * b$, and each $a * b$ must have exactly one result, not two or more.

\subsubsection*{Condition 2: Closed under $*$}
$S$ must be \bld{closed} under $*$. That is,
\[
    \forall~ a,b \in S, \qquad a * b \in S.
\]

\subsubsection{Def. Commutative}
A binary operation $*$ on a set $S$ is commutative if
\[
    \forall~ a,b \in S, \qquad a * b = b * a.
\]

\subsubsection{Def. Associative}
A binary operation $*$ on a set $S$ is associative if
\[
    \forall~ a,b,c \in S, \qquad a * (b * c) = (a * b) * c.
\]

\subsection{Finite Sets}

\subsubsection*{Example}
Let $S = \{a,b,c,d\}$. Define a binary operation $*$ on $S$ using the following table. Complete the table so that $*$ is commutative.
\[
    \begin{array}{c|cccc}
        * & a & b & c & d \\
        \hline
        a & b & d & a & a \\
        b & d & a & c & b \\
        c & a & c & b & b \\
        d & a & b & b & c 
    \end{array}
\]

Note: $*$ is commutative iff the table is symmetric along the main diagonal.

\noindent Is $*$ associative? Why or why not? \bld{No},
\begin{align*}
    a * (b*c) = a*c & = a \\
    (a*b)*c) = d*c & = b
\end{align*}

\subsubsection*{Example}
Suppose that $*$ is associative and commutative operation on a set $S$. Show that $H = \{a \in S : a * a = a\}$ is closed under $*$. Note that the elements of $H$ are called \bld{idenmptents} of the binary operation $*$.
\begin{proof}
    Let $a,b \in H$. Show $a*b \in H$.

    \noindent We know $a*a=a \tand b*b=b$. Show $(a*b)*(a*b)=a*b$.
    \begin{align*}
        LHS & = (a*b)*(a*b) \\
        & = a*(b*a)*b & & \text{since $*$ is associative} \\
        & = a*(a*b)*b & & \text{since $*$ is commutative} \\
        & = (a*a)*(b*b) & & \text{since $*$ is associative} \\
        & = a*b & \\
        & = RHS
    \end{align*}
    Thus, $H$ is closed under $*$.
\end{proof}