\documentclass{article}
\input{../../.latex/article.tex}
\input{../../.latex/macros.tex}

\newcommand{\assignment}{Graded Assignment \#1}

\lhead{MAT 311 Abstract Algebra}
\chead{\assignment}
\rhead{Peter Schaefer}

\begin{document}
\section*{\assignment}

\question{1}{[2 points each] Which of the following are binary operations on the given sets? If it is not an operation, explain why.}
\begin{enumerate}
    \qitem{(a)}{$S=\bb{R}^+ \twith a*b=a\ln b$}{
        A binary operation must be uniquely defined and closed.
        \begin{proof}
            Consider $a = 1 \tand b = \frac{1}{e}$. Both $a \tand b \in \bb{R}^+$, but
            \[
                a*b = a \ln b = 1 \ln \frac{1}{e} = 1 \cdot -1 = -1 \notin \bb{R}^+.
            \]
            Thus $S$ is not closed under $*$, and $*$ cannot be a binary operation.
        \end{proof}
    }
    \qitem{(b)}{$S=\bb{R} \twhere a*b$ is the root of the equation $x^2-a^2b^2=0$}{
        A binary operation must be uniquely defined and closed.
        \begin{proof}
            Consider $a = 2 \tand b = 1$.
            \begin{align*}
                a*b = x^2 - a^2b^2 & = 0 \\
                x^2 - 2^2 1^2 & = \\
                x^2 - 4 & = 0 \\
                (x-2)(x+2) & = 0 \\
                x & = \pm 2
            \end{align*}
            Since there are two solutions, $S$ is not uniquely defined under $*$, and $*$ cannot be a binary operation.
        \end{proof}
    }
\end{enumerate}

\question{2}{[2 points each] Consider the binary operation $*$ defined on $\bb{R}^+$ by $a*b=\frac{ab}{a+b+1}$}
\begin{enumerate}
    \qitem{(a)}{Is $*$ commutative? Explain.}{
        $*$ is commuatative.
        \begin{proof}
            Consider $a*b \tand b*a \tfor a,b \in \bb{R}^+$:
            \[
                a*b = \frac{ab}{a+b+1} = \frac{ba}{b+a+1} = b*a
            \]
            Since $a*b = b*a$ for all $a,b \in \bb{R}^+$, $*$ is commutative.
        \end{proof}
    }
    \qitem{(b)}{Is $*$ associative? Explain.}{
        $*$ is associative.
        \begin{proof}
            Consider $a,b,c \in \bb{R}^+$:
            \begin{align*}
                (a*b)*c & = \frac{ab}{a+b+1}*c = \frac{\frac{ab}{a+b+1}c}{\frac{ab}{a+b+1} +c+1} = \frac{abc}{(a+b+1)(\frac{ab}{a+b+1}+c+1)} \\
                & = \frac{abc}{ab+ac+bc+a+b+c+1} \\ \\
                a*(b*c) & = a*\frac{bc}{b+c+1} = \frac{a\frac{bc}{b+c+1}}{a+\frac{bc}{b+c+1}+1} = \frac{abc}{(b+c+1)(a+\frac{bc}{b+c+1}+1)} \\
                & = \frac{abc}{ab+ac+bc+a+b+c+1}
            \end{align*}
            Thus $(a*b)*c = a*(b*c)$, and $*$ is associative.
        \end{proof}
    }
\end{enumerate}

\question{3}{[3 points] Let $E$ denote the set of all even integers. Prove that $\struct{\bb{Z},+} \simeq \struct{E,+}$.}

\question{4}{[3 points each] Prove that isomorphism is an equivalence relation among binary structures. To do this, you need to prove the following three properties:}
\begin{enumerate}
    \qitem{(a)}{Reflexive: Every binary structure is isomorphic to itself. Hint: let $\struct{S,*}$ be a binary structure and define $\phi: S \rightarrow S$ by $\phi(x)=x$. Prove that $\phi$ is an isomorphism.}{answer}
    \qitem{(b)}{Symmetric: For binary structures $\struct{S_1,*} \tand \struct{S_2,*}$, if $S_1 \simeq S_2$ then $S_2 \simeq S_1$. Hint: assume $\phi: S_1 \rightarrow S_2$ is an isomorphism and prove that $\phi^{-1}: S_2 \rightarrow S_1$ is also an isomorphism.}{answer}
    \qitem{(c)}{Transtiive: For binary structures $\struct{S_1,*},~\struct{S_2,*'}, \tand \struct{S_3,*''}$, if $S_1 \simeq S_2 \tand S_2 \simeq S_3 \tthen S_1 \simeq S_3$. Hind: assume $\phi_1: S_1 \rightarrow S_2 \tand \phi_2: S_2 \rightarrow S_3$ are isomorphisms and prove that $\phi_2 \circ \phi_1: S_1 \rightarrow S_3$ is also an isomorphism.}{answer}
\end{enumerate}

\end{document}