\section{Introduction}

\subsection{Graphs and Graph Models}

\subsubsection*{Graph Definition}
A (simple) \bld{graph} is an ordered pair $(V,E)$ where
\begin{itemize}
    \item $V$ is a nonempty set of objects called "vertices"
    \item $E$ is a set containing some two-subsets of $V$ called "edges". $E$ may be empty.
\end{itemize}

Graphs are often represented pictorially. For example consider
\[
    G = (V,E) \twhere V = \{1,2,3,4,5\} \tand E = \{\{1,4\},\{2,3\},\{2,4\},\{3,4\}\}
\]

\begin{center}
    \begin{tikzpicture}
        \pgfmathsetmacro{\gsize}{2}

        \foreach \rank/\elements/\size in {0/{,1,2}/3, 0.5/{5}/3, 1/{,4,3}/3} {
        \foreach[count = \i] \element in \elements {\node (\element) at (\i * \gsize - \size / 2 * \gsize, -\rank * \gsize) {$\element$};}
        }
        \foreach \j/\l in {1/4,2/3,2/4,3/4} {\draw (\j) -- (\l);}
        \node (name) at (-1,0) {$G:$}; %name
    \end{tikzpicture}
\end{center}
\begin{itemize}
    \item Vertices $1 \tand 4$ are \bld{adjacent} because they are joined by an edge.
    \item Vertex $2$ and edge $2-3$ are \bld{indicent}.
    \item Edges $2-3 \tand 3-4$ are \bld{adjacent}.
\end{itemize}

\subsubsection*{Order Definition}
The \bld{order} of a graph $G$ is $|V(G)|$, or the number of vertices.

\subsubsection*{Size Definition}
The \bld{size} of a graph $G$ is $|E(G)|$, or the number of edges.

The graph $G$ from above has order $5$ and size $4$.

\subsection{Connected Graphs}

\subsubsection*{Subgraph Definition}
Let $G \tand H$ be graphs. $H$ is a \bld{subgraph} of $G$, notated as $H \subseteq G$, if
\[
    V(H) \subseteq V(G) \tand E(H) \subseteq E(G).
\]

\subsubsection*{Proper Subgraph Definition}
$H$ is a \bld{proper subgraph} of $G$ if $H \subseteq G$ and either
\[
    V(H) \subsetneq V(G) \tor E(H) \subsetneq E(G).
\]

\subsubsection*{Spanning Subgraph Definition}
Graph $H$ is a \bld{spanning subgraph} if $H \subseteq G$ and $V(H) = V(G)$.

\subsubsection*{Induced Subgraph Definition}
Graph $H$ is a \bld{induced subgraph} if $H \subseteq G$ and if
\[
    u,v \in V(H) \tand u,v \in E(G) \implies u,v \in E(H).
\]
Essentially, $H$ contains all valid edges it can take from $G$. Notation for \bld{induced subgraph} is
\[
    G[S], \text{ where $S$ is a set of vertices from $G$}.
\]

\subsubsection*{Edge-induced Subgraph Definition}
$G[X]$ is an \bld{edge-induced subgraph} of G if $G[X]$ has edge set $X \subseteq E(G)$ and a vertex set of all vertices incident with at least one edge of $X$. Interesting fact: $G[E(G)]$ removes any isolated vertices.

\subsubsection*{More on Spanning and Induced Subgraphs}
Let $G$ be a graph with vertex $v$ and edge $e$. Then,
\begin{itemize}
    \item $G-e$ is the \itl{spanning subgraph} of $G$ whose edge set is $E(G)-\{e\}$.
          \subitem This definition can be expanded to $G-X \tfor X \subseteq E(G)$.
    \item $G-v$ is the \itl{induced subgraph} of $G$ whose vertex set is $V(G) - \{v\}$ and edge set includes all edges of $G$ except those incident with $v$.
          \subitem This definition can be expanded to $G-U \tfor U \subseteq V(G)$.
\end{itemize}
Let $G$ be a graph, $u,v \in V(G) \tand e = uv \notin E(G)$. Then $G+e$ is the graph with vertex set $V(G)$ and edge set $E(G) \cup \{e\}$. $G$ is a \itl{spanning subgraph} of $G+e$

\subsubsection*{Walk, Trail, Path, Circuit, and Cycle Definitions}
Let $u,v \in V(G)$. A $u-v$ \bld{walk} in $G$ is a sequence of vertices
\[
    (u = v_0, v_1, \ldots, v_k = v)
\]
beginning with $u$, ending with $v$, and consecutive vertices are adjacent.

A \bld{trail} is a walk in which \itl{no edges} are repeated. A \bld{path} is a walk in which \itl{no vertices} are repeated. Every \itl{path} is a \itl{trail} is a \itl{walk}.

A \bld{circuit} is a closed trail of length $\ge 3$. A \bld{cycle} is a circuit with no repeated vertices, except for the first and the last, which are the same. A \bld{$k$-cycle} is a cycle of length $k$. Every \itl{cycle} is a \itl{circuit} is a \itl{walk}.

\subsubsection*{Closed and Open Walks}
A $u-v$ walk with $u=v$ is called a \bld{closed} walk. A $u-v$ walk with $u \neq v$ is called a \bld{open} walk.

\subsubsection*{Walk and Path Theorem}
If $G$ contains a $u-v$ walk of length $\ell$, then $G$ contains a $u-v$ path of length $\leq \ell$.
\begin{proof}
    Let $P = (u=u_0, u_1, \ldots, u_k = v)$ be a $u-v$ walk of smallest length $k \leq \ell$.

    \begin{proof}[Claim]
        $P$ is a $u-v$ path.

        If \bld{not}, then $u_i = u_j$, for some $i \neq j$. Then $(u=u_0, u_1, \ldots, u_i = u_j, \ldots, u_k = v)$ will be a smaller $u-v$ than $P$. This contradicts our assertion that $P$ was the \itl{smallest} walk.
    \end{proof}

    Hence, $P$ is a $u-v$ path of length $k \leq \ell$.
\end{proof}

\subsubsection*{Connectivity Definition}
A graph $G$ is said to be \bld{connected} if $\forall u,v \in V(G)$, $G$ contains a $u-v$ path. If this is not true, i.e. $\exists u,v \in V(G)$ where there is no $u-v$ path, then $G$ is said to be \bld{disconnected}.

\subsubsection*{Component Definition}
A connected subgraph of $G$ that is not a proper subgraph of any other connected subgraph of $G$ is a \bld{component} of $G$. The number of components of a graph $G$ is denoted by $k(G)$. A graph $G$ is connected if and only if $k(G) = 1$. Additionally, a graph is the union of its components.

\subsubsection*{Components and Equivalence Relations Theorem}
Define a relation $R$ on the $V(G)$ so that $uRv$ if $G$ contains a $u-v$ walk. Then $R$ is an equivalence relation.
\begin{proof}
    An equivalence relation must be reflexive, symmetric, and transitive.
    \begin{enumerate}
        \item Reflexive: $\forall~u \in V(G)$, $(u)$ is a $u-u$ walk, so $uRu$.
        \item Symmetric: Suppose $uRv$. There is a $u-v$ walk $(u=u_0, u_1, \ldots, u_n = v)$. Then reversing the walk gives the $v-u$ walk $(v=u_n, \ldots, u_1, u_0 = u)$. So $vRu$.
        \item Transitive: Suppose $uRv \tand vRw$. There is a $u-v$ walk $(u=u_0, u_1, \ldots, u_n = v)$, and $v-w$ walk $(v=v_0, v_1, \ldots, v_m = w)$. Then there is a $u-w$ walk $(u=u_0, u_1, \ldots, u_n = v = v_0, \ldots, v_m = w)$. So $uRw$.
    \end{enumerate}
\end{proof}
What are the equivalence classes?
\begin{align*}
    [u] & = \{v \in V(G) ~|~ uRv\}                                 \\
        & = \{v \in V(G) ~|~ \text{there is a $u-v$ walk in $G$}\} \\
        & = \text{the connected component containing $u$}
\end{align*}

\subsubsection*{Subtractive Connectivity Theorem (weak)}
Let $G$ be a graph of order $\geq 3$. If $\exists~ u,v \in V(G)$ such that $G-u$ and $G-v$ are connected, then $G$ is connected.
\begin{proof}
    Suppose $G$ has order at least 3 and $\exists~ u,v \in V(G)$ such that $G-u \tand G-v$ are connected. Let $x,y \in V(G)$.
    \begin{enumerate}[start=1,label={\bfseries Case \arabic*:},leftmargin=0.75in]
        \item $\{x,y\} \neq \{u,v\}$, meaning at least one is different. Say (WLOG) $u \notin \{x,y\}$.

              Then $x,y \in V(G-u)$, which is connected, contains an $x-y$ walk.
        \item $\{x,y\} = \{u,v\}$. Say (WLOG) $x=u \tand y=v$.

              Consider $z \in V(G-u) \tand z \in V(G-v)$.

              Then $x,z \in V(G-v)$ contains a $x-z$ walk $(x=x_0, \ldots, x_n = z)$, since it is connected.

              Then $z,y \in V(G-u)$ contains a $z-y$ walk $(z=z_0, \ldots, z_m = y)$, since it is connected.

              Now consider $(x=x_0, \ldots, x_n = z = z_0, \ldots, z_m = y))$. This is an $x-y$ walk in $G$.
    \end{enumerate}
\end{proof}

\subsubsection*{Distance, Geodesic, Diameter, and Girth Definitions}
The \bld{distance} between vertices $u \tand v$, denoted as $d(u,v) \tor d_G(u,v)$ is the smallest length of any $u-v$ path in $G$. If $u \tand v$ are in different components, then $d(u,v)$ is undefined.

A $u-v$ path of shortest length $d(u,v)$ is called a \bld{geodesic}. The \bld{diameter} of a connected graph $G$, denoted as $\diam(G)$, is the largest \itl{geodesic} between any two vertices of $G$. The \bld{girth} of a connected graph $G$ is the length of the shortest cycle in $G$.

\subsubsection*{Subtractive Connectivity Theorem (strong)}
Let $G$ be a graph of order $\geq 3$. Then $G$ is connected if and only if $\exists~ u,v \in V(G)$ such that $G-u$ and $G-v$ are connected.
\begin{proof}[Proof $\implies$]
    This direction is already proven by the weak version of this theorem.
\end{proof}
\begin{proof}[Proof $\impliedby$]
    Suppose $G$ is connected. Then $\exists~ u,v \in V(G)$ such that $d(u,v) = \diam(G)$.

    Suppose to the contrary, WLOG, that $G-v$ is disconnected. Then $exists~x,y \in V(G-v)$ such that there is no $x-y$ walk  in $G-v$. But $G$ is connected, so there exist $x-u \tand u-y$ paths in $G$.

    Let $P'$ be an $x-u$ geodesic in $G$ and $P''$ be a $u-y$ geodesic in $G$.

    $v$ cannot be on $P' \tor P''$ because if it was, then $d(x,u) > d(u,v) \tor d(y,u) > d(u,v)$, violating our assertion that $d(u,v) = \diam(G)$.

    Then $P'P''$ $x-u-y$ is an $x-y$ walk in $G-v$.

    This contradicts our selection that $x \tand y$ do not have a walk in $G-v$. Hence $G-v \tand G-u$ are connected.
\end{proof}

\subsection{Common Classes of Graphs}
\begin{center}
    \begin{tabular}{|c|c|c|c|}
        \hline
        \bld{Name}         & \bld{Symbol} & \bld{Order} & \bld{Size}     \\
        \hline
        Path               & $P_n$        & $n$         & $n-1$          \\
        Cycle              & $C_n$        & $n \geq 3$  & $n$            \\
        Complete           & $K_n$        & $n$         & $\binom{n}{2}$ \\
        Complete Bipartite & $K_{s,t}$    & $s+t$       & $s \cdot t$    \\
        \hline
    \end{tabular}
\end{center}

\subsubsection*{Bipartite Graph Definition}
$G$ is bipartite if $V(G)$ can be partitioned into partite sets $U \tand W$ so that every edge joins a vertex of $U$ and a vertex of $W$.

\subsubsection*{Odd Cycle and Bipartite-ness Theorem}
$G$ is bipartite if and only if $G$ contains no odd cycles.
\begin{proof}[Proof $\implies$]
    Via contradiction, suppose $G$ contains an odd cycle $C$ and $G$ is bipartite with partite sets $U \tand V$.
    \[
        C = (u_1,u_2,\ldots,u_{2n},u_{2n+1},u_1)
    \]
    Without loss of generality, assume $u_1 \in U$. Then $u_1,u_3,u_5,\ldots,u_{2n+1} \in U$. But $u_{2n+1}$ and $u_1$ are adjacent and both are in $U$.
\end{proof}
\begin{proof}[Proof $\impliedby$]
    Suppose $G$ has no odd cycles. Assume $G$ is connected, or a connected component of a larger graph. Let $u \in V(G)$.

    Define
    \begin{align*}
        U & = \{v \in V(G) ~|~ d(u,v) ~\text{is even.}\} \\
        W & = \{v \in V(G) ~|~ d(u,v) ~\text{is odd.}\}
    \end{align*}
    This is a partition of $V(G)$ and $u \in U$ as $d(u,u) = 0$, which is even.

    Prove every edge join a vertex in $U$ and a vertex in $W$.
    \begin{proof}[Subproof]
        Suppose, to the contrary, there are $v,w \in W$ with $vw \in E(G)$.

        Note that $d(u,v) \tand d(u,w)$ are odd. Let
        \begin{align*}
            d(u,v) & = 2s+1                          \\
            d(u,w) & = 2t+1 & \tfor s,t \in \bb{Z}^+
        \end{align*}
        Consider
        \begin{align*}
            P'  & = (u=v_0,v_1,\ldots,v_{2s+1} = v) \\
            P'' & = (u=w_0,w_1,\ldots,w_{2t+1} = w)
        \end{align*}
        $P' \tand P''$ have $u$ in common, and maybe other vertices as well. Let $x$ be the last vertex in common.

        So $x=v_i$ for some $0 \leq i \leq 2s+1 \tand d(u,v_i) = i$.

        So $x=w_i$ for some $0 \leq i \leq 2t+1 \tand d(u,w_i) = i$.
        \begin{align*}
            P'  & = (u=v_0,v_1,\ldots,v_i,\ldots,v_{2s+1} = v) \\
            P'' & = (u=w_0,w_1,\ldots,w_i,\ldots,w_{2t+1} = w)
        \end{align*}
        Since $vw \in E(G)$, we have a cycle $C = (v_i,v_{i+1},\ldots,v_{2s+1},w_{2t+1},w_{2t},\ldots,w_i=v_i)$ of length
        \[
            \underbrace{2s+1-i+1}_\text{top row} + \underbrace{2t+1-i}_\text{bottom row} = 2s + 2t - 2i + 1 = 2(s+t-i)+1
        \]
        So $C$ is an odd cycle, which contradicts our assertion that $G$ has \uln{no} odd cycles.
    \end{proof}
\end{proof}

\subsubsection*{K-partite Definition}
$G$ is a \bld{$k$-partite} graph if $V(G)$ can be partitioned into partite sets $U_1, \ldots, U_k$ so that every edge joins a vertex from $U_i$ and a vertex of $U_j$ where $i \neq j$.

\subsection*{Constructing New Graphs from Old Graphs}
\subsubsection*{Disjoint Union}
For two graphs $G \tand H$, $G \cup H$ is defined as...
\begin{align*}
    V(G \cup H) & = V(G) \cup V(H) \\
    E(G \cup H) & = E(G) \cup E(H)
\end{align*}

\subsubsection*{Complement}
For one graph $G$, $\overline{G}$ is defined as...
\begin{align*}
    V(\overline{G}) & = V(G)                                            \\
    E(\overline{G}) & = \{uv | u,v \in V(G), u \neq v, uv \notin E(G)\}
\end{align*}

\subsubsection*{Join}
For two graph, $G \tand H$, $G + H$ is defined as...
\begin{center}
    Start with $G \cup H$ and draw all edges join a vertex of $G$ and a vertex of $H$
\end{center}

\subsubsection*{Cartesian Product}
For two graphs, $G \tand H$, $G \times H$ is defined as...
\begin{align*}
    V(G \times H) & = \{(u,v) | u \in V(G) \tand v \in V(H)\}                 \\
    (u,v)-(x,y)   & \tif u = x \tand vy \in E(H) \lor v = y \tand ux \in E(G)
\end{align*}
A cartesian product between two graphs has the practical effect of duplicating one graph, and connecting the duplicates in the way of the other graph.

\subsubsection*{Complement Connectivity Theorem}
If $G$ is disconnected, then $\overline{G}$ is connected.
\begin{proof}
    Let $u,v \in V(\overline{G})$.
    \begin{enumerate}[start=1,label={\bfseries Case \arabic*:},leftmargin=0.75in]
        \item If $u,v$ are in different components of $G$, then $u,v \in E(\comp{G})$, so $(u,v)$ is a walk in $\comp{G}$.
        \item If $u,v$ are in the same component of $G$, then $\exists~ w \in V(G)$ in a different component. So $uw, wv \in E(\comp{G})$.

              Hence $(u,w,v)$ is a $u-v$ walk.
    \end{enumerate}
\end{proof}

\subsection{Multigraphs and Digraphs}

\subsubsection*{Multigraph Definition}
A \bld{multigraph} is a graph where a pair of vertices may be joined by any finite number of edges.
\begin{itemize}
    \item Multiple edges: OK
    \item Loops: NOT OK
\end{itemize}

\subsubsection*{Pseudograph Definition}
A \bld{pseudograph} is a \itl{multigraph} where loops are allowed
\begin{itemize}
    \item Multiple edges: OK
    \item Loops: OK
\end{itemize}

\subsubsection*{Digraph Definition}
A \bld{directed graph} is a graph where $E(G)$ is a set of ordered pairs (rather than sets) of distinct vertices called directed edges, or arcs.

\subsubsection*{Oriented Graph Definition}
An \bld{oriented graph} is a \itl{digraph} in $\forall u,v \in V(G)$, $(u,v) \tand (v,u)$ are not \underbar{both} edges.